Left-linear rules in adhesive categories

### 2024-02-12 icalp/fusione.tex
examples

@@

\begin{example}
  The result above fails when considering left-linear
  derivations. This is readily seen by considering derivations
  $\der{D}$ in Fig.~\ref{fi:derD} and $\der{D}'$ in
  Fig.~\ref{fi:derD1}. We already observed that
  $\der{D} \not\equiv_a \der{D}'$generale}
\end{example}

@@

Instead, if in $r_0\, r_1\, r_2$ we have $r_1$ and $r_2$ are
independent, and we switch $r_0$ to the last position, obtaining a
derivation $r_1\, r_2\, r_0$, the steps $r_1$ and $r_2$ could be no
longer independent, unless an additional condition is
satisfied. Intuitively this is due to the fact that, dealing with
left-linear rewriting systems introduces disjunctive forms of
causality. Hence, if in $r_0\, r_1\, r_2$ we have that $r_0, r_1$ are
disjunctive causes of $r_2$, at least one of the two is needed to
enable the application of $r_2$ and this explain why in
$r_1\, r_2\, r_0$ we have that $r_1$ and $r_2$ are no longer
independent. \todo{P: da dire meglio e Ci vorrebbe un esempio}

>>>>

    \caption{Rules of the rewriting system of the running example.}
    \label{fi:rules}
  \end{figure}

@@

\begin{example}
  The result above fails when considering left-linear
  derivations. This is readily seen by considering derivations
  $\der{D}$ in Fig.~\ref{fi:derD} and $\der{D}'$ in
  Fig.~\ref{fi:derD1}. We already observed that
  $\der{D} \not\equiv_a \der{D}'$\todo{P: osservato?}
  %
  Still, it is easy to see that they are related by a consistent
  permutation, which is the identity on productions. The colimit of
  both derivations is a single node with a self loop, i.e., the
  terminal object in $\cat{Graph}$, hence all commutation requirements
  are trivially satisfied.
\end{example}

@@

Instead, if in $r_0\, r_1\, r_2$ we have $r_1$ and $r_2$ are
independent, and we switch $r_0$ to the last position, obtaining a
derivation $r_1\, r_2\, r_0$, the steps $r_1$ and $r_2$ could be no
longer independent, unless an additional condition is
satisfied. Intuitively this is due to the fact that, dealing with
left-linear rewriting systems introduces disjunctive forms of
causality. Hence, if in $r_0\, r_1\, r_2$ we have that $r_0, r_1$ are
disjunctive causes of $r_2$, at least one of the two is needed to
enable the application of $r_2$ and this explain why in
$r_1\, r_2\, r_0$ we have that $r_1$ and $r_2$ are no longer
independent.

\begin{example}
  Consider the rewriting system used as running example (rules in
  Fig.~\ref{fi:rules}. Assume to add a fourth rule $\rho_4$ as in
  Fig.~\ref{fi:rule-cas} and consider the derivation.
\end{example}
